# Part 2: Becoming a Proud Bayesian

For our finalized Bayesian model, we intend to build predictive models for universities and liberal arts colleges. Then we'll compare the similarities and differences of the two models. (i.e. is some factor important for universities and less so for liberal arts colleges?)

## University Model

After deciding on the variables of interest (size of undergraduate student body, average SAT score, admission rate, and location), we took a step-by-step approach to create the model. In addition to determining the variables we needed, we also realized that we had to change our model type. After repeatedly getting negative values from our Normal model, we decided on a Poisson model. This would ensure positive outputs and we were hopeful that it would fit our data better.

### Get some intuition

For some intuition, we built some bi-variate visualizations that helped us understand the individual trends between ranking and our quantitative variables: 

```{r message=FALSE, warning=FALSE, echo=FALSE}
gg1 <- ggplot(fullUniversity, aes(x = UGDS_1617, y = as.numeric(Y2018))) +
  geom_point() +
  geom_smooth(method = "glm") +
  xlab("Number of Undergraduate Students in 2016-17") +
  ylab("School Ranking 2018") + ggtitle("Universities") + ylim(1,120)

gg2 <- ggplot(fullUniversity, aes(x = SAT_AVG_1617, y = as.numeric(Y2018))) +
  geom_point() +
  geom_smooth(method = "glm") +
  xlab("Average SAT Score in 2016-17") +
  ylab("School Ranking 2018") + ylim(1, 120)
gg3 <- ggplot(fullUniversity, aes(x = ADM_RATE_1617, y = as.numeric(Y2018))) +
  geom_point() +
  geom_smooth(method = "glm") +
  xlab("Admission Rate in 2016-17") +
  ylab("School Ranking 2018") + ylim(1,120)
grid.arrange(gg1, gg2, gg3, nrow = 2)
```

After looking at the following plots and fitting them with generalized linear models (since Poisson models are on the log-scale), we saw generally linear relationships between rank and SAT score and admission rate. There was a pretty weak linear relationship between ranking and number of undergrads. However, having seen these plots and fit linear models to them, we were able to still move forward with our generalized linear model.

### Some addition intuition

For more intuition, we start with more vague priors and slightly adjust the priors based on the summary table. It’s important to note that the priors cannot be set too close to the means of the summary table in case of **overfitting**.

### Building the model


```{r}
university_model_3 <- "model{  
    # Data: observations
    for(i in 1:length(y)) {
        y[i] ~ dpois(lambda[i])
        log(lambda[i]) = beta0 + beta1*x1[i] + beta2[z[i]] + beta3*x3[i] + beta4*x4[i]
    }
        # Data: subjects
        beta0 ~ dnorm(8, 1)
        beta1 ~ dnorm(0, 10000)
        beta2[1] <- 0 
        beta2[2] ~ dnorm(0, 0.01)
        beta2[3] ~ dnorm(0, 0.01)
        beta3 ~ dnorm(0, 10)
        beta4 ~ dnorm(0, 0.01)

}"

# COMPILE

y <- as.numeric(fullUniversity$Y2018)

model_data9 <- data.frame(y, x1 = fullUniversity$UGDS_1617, z = as.numeric(fullUniversity$LOCALE_collapse), x3 = fullUniversity$SAT_AVG_1617, x4 = fullUniversity$ADM_RATE_1617)

model_data9 <- na.omit(model_data9)

university_jags_3 <- jags.model(textConnection(university_model_3), 
    data = list(y = model_data9$y, x1 = model_data9$x1, z = factor(model_data9$z), x3 = model_data9$x3, x4 = model_data9$x4),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

# SIMULATE the model
university_sim_3 <- coda.samples(university_jags_3,
    variable.names = c("beta0","beta1","beta2","beta3","beta4"),
    n.iter = 10000)

# STORE the chains in a data frame
university_chains_3 <- data.frame(university_sim_3[[1]])
```


### Model summary

```{r}
summary(university_sim_3)
```


### Posterior inference

For an unknown university located in the city (beta2 = 0, since city is the reference level) with 10000 undergraduates, student mean SAT score of 1400 and an admission rate of $25\%$ (e.g. Gvictor University), we could predict its ranking from our rjags simulation.

```{r}
university_chains_3 <- university_chains_3 %>%
  mutate(ranking_new = rpois(10000, lambda = exp(beta0 + beta1 * 10000 + beta3 * 1400 + beta4 * 0.25)))

university_chains_3 %>%
  summarize(quantile(ranking_new,0.025),quantile(ranking_new,0.975))
```

The interval is more reflective of the intuitive estimate of Gvictor University's ranking. We get a reasonable range of values (1, 120), and we would predict that a school with a moderately selective admissions rate (25%) and an SAT score average of 1400, would fall into a "Tier 2" range (approximately between 25 and 50).

## Liberal Arts Colleges

### Get some intuition

For some intuition, we again built some bi-variate visualizations that helped us understand the individual trends between ranking and our quantitative variables: 

```{r message=FALSE, warning=FALSE, echo=FALSE}
gg4 <- ggplot(full_LiberalArts, aes(x = UGDS_1617, y = as.numeric(Y2018))) +
  geom_point() +
  geom_smooth(method = "glm") +
  xlab("Number of Undergraduate Students in 2016-17") +
  ylab("School Ranking 2018") + ggtitle("Liberal Arts Colleges") + ylim(1,150)

gg5 <- ggplot(full_LiberalArts, aes(x = SAT_AVG_1617, y = as.numeric(Y2018))) +
  geom_point() +
  geom_smooth(method = "glm") +
  xlab("Average SAT Score in 2016-17") +
  ylab("School Ranking 2018") + ylim(1, 150)
gg6 <- ggplot(full_LiberalArts, aes(x = ADM_RATE_1617, y = as.numeric(Y2018))) +
  geom_point() +
  geom_smooth(method = "glm") +
  xlab("Admission Rate in 2016-17") +
  ylab("School Ranking 2018") + ylim(1,150)
grid.arrange(gg4, gg5, gg6, nrow = 2)
```
After looking at the following plots and fitting them with generalized linear models, we saw the strongest linear relationship between average SAT score and ranking. There was also a fairly strong linear relationship between admission rate and ranking. The relationship between ranking and number of undergrads remain quite weak for liberal arts colleges, but still existent. So, having seen these plots and fit linear models to them, we were confident in moving forward.

### Building the model


```{r}
liberal_model_3 <- "model{  
    # Data: observations
    for(i in 1:length(y)) {
        y[i] ~ dpois(lambda[i])
        log(lambda[i]) = beta0 + beta1*x1[i] + beta2[z[i]] + beta3*x3[i] + beta4*x4[i]
    }
    # Data: subjects
    beta0 ~ dnorm(0, 0.0001)
    beta1 ~ dnorm(0, 10000)
    beta2[1] <- 0 
    beta2[2] ~ dnorm(0, 0.04)
    beta2[3] ~ dnorm(-10, 0.04)
    beta2[4] ~ dnorm(-15, 0.01)
    beta3 ~ dnorm(-0.3, 100)
    beta4 ~ dnorm(36, 10)

}"

# COMPILE

y <- as.numeric(full_LiberalArts$Y2018)

model_data8 <- data.frame(y, x1 = as.numeric(full_LiberalArts$UGDS_1617), z = as.numeric(full_LiberalArts$LOCALE_collapse_lac1), x3 = as.numeric(full_LiberalArts$SAT_AVG_1617), x4 = as.numeric(full_LiberalArts$ADM_RATE_1617))

model_data8 <- na.omit(model_data8)

liberal_jags_3 <- jags.model(textConnection(liberal_model_3), 
    data = list(y = model_data8$y, x1 = model_data8$x1, z = factor(model_data8$z), x3 = model_data8$x3, x4 = model_data8$x4),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

# SIMULATE the model
liberal_sim_3 <- coda.samples(liberal_jags_3,
    variable.names = c("beta0","beta1","beta2","beta3","beta4"),
    n.iter = 10000)

# STORE the chains in a data frame
liberal_chains_3 <- data.frame(liberal_sim_3[[1]])
```

### Model summary

```{r}
summary(liberal_sim_3)
```


### Posterior inference

For an unknown liberal arts college in a city setting (beta2 = 0, since city is the reference level) with 2000 undergraduates, student mean SAT score of 1400 and an admission rate of $25\%$ (e.g. Hvictor College), we could predict its ranking from our rjags simulation.

```{r}
liberal_chains_3 <- liberal_chains_3 %>%
  mutate(ranking_new = rpois(10000, exp(beta0 + beta1 * 2000 + beta3 * 1400 + beta4 * 0.25)))

liberal_chains_3 %>%
  summarize(quantile(ranking_new,0.025),quantile(ranking_new,0.975))
```

This is **totally reasonable**. As mentioned earlier, liberal arts colleges tend to take a more holistic approach to their admissions processes and often have a smaller pool of applicants that is pretty self-seleting. As such, they tend to put less weight on standardized testing. As such, we can expect that such a school might have a pretty broad range of rankings it could receive. The interval reflects the intuitive estimate of Hvictor College's ranking.

## Comparison & Discussion

Compared to the last attempts, our final model gives reasonable predictions, which is a lot of progress! However, the 95% credible interval doesn’t provide as accurate an interval as desired.

By creating different imaginary schools, we are able to summarize the most important factors for universities and colleges respectively. We find that universities value SAT score as a more decisive indicator of both students’ admission and their own ranking. We also supported our intuition that liberal arts college tend distribute their weight among many factors when determining admission.
